---
title: 'Data handling and EDA: For Presentation'
author: "Tomohiro Sasaki"
date: "2020年9月30日"
output: html_document
editor_options: 
  chunk_output_type: console
---


## Data handling

Welcome to **data handling** session!  
Tutors try to provide accurate code. However, incorrect one may be included. Please use it at your own risk.  
This document was built by `r R.Version()$version.string`.  

```{r setup, include=FALSE}
#install.packages("knitr")
#install.packages("tidyverse")
#install.packages("dplyr")

library(knitr)
library(tidyverse)
library(dplyr)

Sys.setlocale('LC_ALL','C')

knitr::opts_chunk$set(echo = TRUE)
```

## Set up

Before handling the dataset, set some libraries.

```{r settings}
#path <- "F:/Desktop/Rpmx/R-for-PMx"               # enviroment - 1
path <- "C:/Users/Sasaki.Tomohiro/Desktop/R-for-PMx"  # enviroment - 2
# default path <- "/cloud/project/R-for-PMx-master"
```


## Import data

Let's import nonmem dataset!

```{r import_and_check}
nm_data <- read_csv(paste0(path, "/Data/PSP4-8-748-s012.csv")) ## csv fileの読み込み

nm_data <- read_csv(paste0(path, "/Data/PSP4-8-748-s012.csv"), na=".",
                    col_types = cols(C     = col_character(),
                                     TREAT = col_character())
                    )

print(head(nm_data, 10))    ## 最初の10行の表示                   

# print(nrow(nm_data))
# print(ncol(nm_data))
# print(names(nm_data))

#　列数, 行数, 列名の表示
nm_data %>% nrow() %>% print()
nm_data %>% ncol() %>% print()
nm_data %>% names() %>% print()

# スプレッドシートでの表示
# View(nm_data)
```

## Exploratory data analysis

Let's see the dataset by summary statistics and visualization!

### Generate Summary statistics

```{r Sum-stats}
## group_by, filter, summarize, slice, n, count
# 薬剤ごとの濃度ポイント数（conc>0）の算出
nm_data %>% group_by(DRUG) %>% filter(CONC > 0) %>% summarise(Nobs = n())
# 薬剤ごと, PN(試験)ごとの濃度ポイント数（conc>0）の算出
nm_data %>% group_by(DRUG, PN) %>% filter(CONC > 0) %>% summarise(Nobs = n())

# IDごとに1行のデータを作成
sl_nm_data <- nm_data %>% group_by(ID) %>% slice(1) %>% ungroup()

# ID数の表示
sl_nm_data %>% nrow() %>% print()
# PN(試験)ごとのID数の表示
sl_nm_data %>% count(PN)

sl_nm_data %>% summarise(Number = n(),
                         meanCRCL = mean(CRCL),
                         meanWT = mean(WT),
                         meanAGE = mean(AGE))

sl_nm_data %>% group_by(PN) %>% summarise(N = n(),
                                          meanCRCL = mean(CRCL) %>% round(1),
                                          meanWT = mean(WT) %>% round(1),
                                          meanAGE = mean(AGE) %>% round(1)) %>% knitr::kable()

sl_nm_data %>% group_by(PN) %>% summarise(Number = n(),
                                          meanCRCL = mean(CRCL) %>% round(1),
                                          meanWT = mean(WT) %>% round(1),
                                          meanAGE = mean(AGE) %>% round(1),
                                          sdCRCL = sd(CRCL) %>% round(1),
                                          sdWT = sd(WT) %>% round(1),
                                          sdAGE = sd(AGE) %>% round(1)) %>% 
  mutate(CRCL = paste(meanCRCL, sdCRCL, sep = "+/-")) %>% 
  mutate(WT = paste(meanWT, sdWT, sep = "+/-")) %>% 
  mutate(AGE = paste(meanAGE, sdAGE, sep = "+/-")) %>%
  select(PN, Number, CRCL, WT, AGE) %>% 
  rename(Protocol_ID = PN) %>% 
  knitr::kable(align = "c")

```

```{r Ex-1}
## Read dataset
nm_data2 <- read_csv(paste0(path, "/Data/PSP4-8-748-s012.csv"), na=".",
                     col_types = cols(C     = col_character(),
                                      TREAT = col_character() 
                                      ))

## Number of Subject by PN
nm_data2 %>% group_by(ID) %>% slice(1) %>% group_by(PN) %>% count() 

## Number of MK7655 Vaid Concentration Record by PN
nm_data2 %>% filter(MDV == 0 & DRUG == 2) %>% group_by(PN) %>% count() 

## Number of Subject and MK7655 Vaid Concentration Record by PN
nm_data2 %>% group_by(PN) %>% summarise(N_Sub  = length(unique(ID)),
                                        N_Conc = length(CONC[MDV == 0 & DRUG == 2])) 

## Desc. Stats. Grouped by PN and Dose
nm_data2 %>% filter(PN <= 2) %>% group_by(PN, DOSE) %>% summarise(Age_Mean  = mean(AGE),
                                                                  WT_Mean   = mean(WT),
                                                                  CRCL_Mean = mean(CRCL))

```

```{r Ex-2}
## Read dataset
nm_data2 <- read_csv(paste0(path, "/Data/PSP4-8-748-s012.csv"), na=".",
                     col_types = cols(C     = col_character(),
                                      TREAT = col_character() 
                                      ))
back_info <- read_csv(paste0(path, "/Data/back_info.csv"), na=".")

## Merge dataset & count ID by CYP2D6
nm_data2 %>% left_join(back_info, by="ID") %>% group_by(ID) %>% slice(1) %>% group_by(CYP2D6) %>% count()

```

```{r Join-Examples}
tibA <- tibble(ID   = c(rep(c(1, 2), each=3)),
               TIME = c(0, 1, 2, 0, 2, 4))

tibB <- tibble(id  = rep(c(1, 2), each=2), 
               tad = c(0, 1, 0, 3), 
               PD  = c(100, 90, 100, 70))

tibA %>% left_join(tibB, by=c("ID"="id", "TIME"="tad"))
tibA %>% right_join(tibB, by=c("ID"="id", "TIME"="tad"))
tibA %>% inner_join(tibB, by=c("ID"="id", "TIME"="tad"))
tibA %>% full_join(tibB, by=c("ID"="id", "TIME"="tad"))

tibC <-  tibble(ID   = c(1, 2, 1, 1, 2),
                TIME = c(0, 0, 0, 2, 2),
                MDV  = c(0, 1, 1, 0, 0))

tibC %>% arrange(ID, TIME, desc(MDV))

```

```{r Other-Examples}
nm_data_e <- expand.grid(ID = 1:10, TIME=1:24) %>% as_tibble()

nm_data %>% group_by(ID) %>% slice(1) %>% ungroup() %>% sample_n(1000, replace = TRUE)

mut_dat <-  tibble(ID = 1:3,
                   HT = c(160, 170, 180),
                   SEX = c(2, 2, 1))

mut_dat %>% mutate(IBW = 22*(HT/100)^2)
mut_dat %>% mutate(IBW = ifelse(SEX == 1, 50 + 0.91*(HT-152.4), 45.5 + 0.91*(HT-152.4)))

exp_dat <- expand.grid(ID = 1:3,
                       TIME = 0:2)

```

```{r back_info}

Pheno2D6 <- c(rep(1, 500), rep(2, 350), rep(3, 50), rep(4, 100))

back_info <- nm_data %>% group_by(ID) %>% slice(1) %>% ungroup()

back_info <- back_info %>% select(ID) %>% mutate(CYP2D6 = sample(Pheno2D6, nrow(back_info), replace = TRUE))

write.table(back_info, paste("Data", "/", "back_info.csv", sep=""), sep=",", row.name=F, na=".", quote=F, fileEncoding = "ASCII")
   
```

### Visualize data

```{r}
p <- ggplot(sl_nm_data %>% select(PN, CRCL, WT, AGE))
p <- p + geom_point(aes(x = WT, y = CRCL))
p

p <- p + facet_wrap(~ PN)
p

p <- ggplot(sl_nm_data %>% select(PN, CRCL, WT, AGE, HLTH))
p <- p + geom_point(aes(x = WT, y = CRCL, fill = factor(HLTH)), shape = 21, alpha = 0.5)
p <- p + theme_bw()
p <- p + labs(x = "Body weight (kg)", y = "Creatine clearance (mL/min)", fill = "Healthy volunteer")
p <- p + scale_x_continuous(limits = c(25, NA))
p <- p + scale_y_continuous(limits = c(0, NA))
p <- p + theme(legend.position = "top")
p

```

```{r ggplot}
 
## Data to plot
nm_data_p <- nm_data %>% filter(PN==19 & DRUG==2 & MDV==0 & PERD == 1)

## ggplot
nm_data_p  %>% ggplot(aes(x=TIME, y=CONC, group=ID, colour=MALE))
ggsave("sample/1_data_handling/ggplot.png", width=10, height = 8)

## geom_point
nm_data_p  %>% ggplot(aes(x=TIME, y=CONC, group=ID, colour=MALE)) + geom_point()
ggsave("sample/1_data_handling/geom_point.png", width=10, height = 8)
nm_data_p  %>% ggplot(aes(x=TIME, y=CONC, group=ID, colour=factor(MALE))) + geom_point()
ggsave("sample/1_data_handling/geom_point_fMALE.png", width=10, height = 8)

## geom_line
nm_data_p  %>% ggplot(aes(x=TIME, y=CONC, group=ID, colour=factor(MALE))) + geom_line()
ggsave("sample/1_data_handling/geom_line_fMALE.png", width=10, height = 8)


## stat_summary
ggplot(data=nm_data_p, aes(x=TIME, y=CONC, colour=factor(MALE))) + stat_summary()
ggplot(data=nm_data_p, aes(x=TIME, y=CONC, colour=factor(MALE))) + stat_summary(fun=median) + 
stat_summary(fun=median, geom="line")
ggsave("sample/1_data_handling/stat_summary.png", width=10, height = 8)

## stat_smooth
ggplot(data=nm_data_p, aes(x=TIME, y=CONC, colour=factor(MALE))) + stat_smooth()
ggsave("sample/1_data_handling/stat_smooth.png", width=10, height = 8)

## facet_wrap
ggplot(data=nm_data_p, aes(x=TIME, y=CONC, colour=factor(MALE))) + geom_line() + facet_wrap(~ID, scales = "free")
ggsave("sample/1_data_handling/facet_wrap.png", width=10, height = 8)

## Data to plot both drug=1&2
nm_data_p2 <- nm_data %>% filter(PN==19 & MDV==0 & PERD == 1)

## facet_grid
ggplot(data=nm_data_p2, aes(x=TIME, y=CONC, colour=factor(MALE))) + geom_line(aes(group=ID)) + facet_grid(DRUG~MALE, scales="free")
ggsave("sample/1_data_handling/facet_grid.png", width=10, height = 8)

```

```{r ggplot-2}
## Data to plot
nm_data_b <- nm_data %>% group_by(ID) %>% slice(1) %>% ungroup()

## geom_boxplot
nm_data_b %>% ggplot() + geom_boxplot(aes(x=factor(PN), y=CRCL))
ggsave("sample/1_data_handling/geom_boxplot.png", width=10, height = 8)

## geom_histogram
nm_data_b %>% ggplot() + geom_histogram(aes(x=CRCL, fill=factor(PN)), position = "identity", alpha = 0.5, binwidth = 10)
ggsave("sample/1_data_handling/geom_boxplot_PN.png", width=10, height = 8)

## Data to plot both drug=1&2
nm_data_p2 <- nm_data %>% filter(PN==19 & MDV==0 & PERD == 1)

## facet_grid
ggplot(data=nm_data_p2, aes(x=TIME, y=CONC, colour=factor(MALE))) + geom_line(aes(group=ID)) + facet_grid(DRUG~MALE, scales="free") + 
  xlab("Time after dose(hr)") + ylab("Plasma Concentration (ng/mL)") +
  scale_x_continuous(breaks = seq(0, 10, by=2)) + 
  scale_colour_discrete(name = "Sex (0=Female, 1=Male)") +
  ggtitle("Plasma Concentration Profiles by Sex", subtitle = "Only PN=19 data used") +
  labs(caption = paste0("Y panel: 1=Imipenem, 2=MK7655, X panel: 0=Female, 1=Male\n",
                        "Source: ", path, "/Data/PSP4-8-748-s012.csv"
                        )
       ) +
  theme_bw() +
  theme(axis.text= element_text(size=10),
        legend.title = element_text(size=10),
        legend.text = element_text(size=10),
        legend.key.size = unit(0.3, "cm"),
        plot.title= element_text(size=15, face="bold"),
        plot.subtitle = element_text(size=10),
        plot.caption = element_text(size=8, colour="black", hjust = 1.0),
        strip.text=element_text(size=10),
        legend.position = "top")

ggsave("sample/1_data_handling/facet_grid_formatting.png", width=10, height = 8)
```

### For Presentation

```{r Demo-Stats, eval=FALSE}
library(flextable)

## Continuous Covariates

nm_stats <- nm_data %>% filter(PN <= 4) %>% group_by(ID) %>% slice(1) %>%
  mutate_at(c("WT", "AGE", "CRCL", "HT"), ~as.double(gsub(., pattern=-99, replacement = NA))) %>% 
  group_by(PN) %>% 
  summarise_at(c("WT", "AGE", "CRCL", "HT"),
              list(~length(which(!is.na(.))), ~min(., na.rm = TRUE), 
                   ~max(., na.rm = TRUE), ~mean(., na.rm = TRUE), 
                   ~sd(., na.rm = TRUE), ~median(., na.rm = TRUE))) %>% 
  mutate_if(is.double, ~signif(., 3)) %>% 
  gather(val=Value, key=Key, -PN) %>% 
  separate(Key, c("Covariate", "Stats"), "_") %>% 
  spread(key=Stats, value=Value) %>% 
  mutate(Covariate = factor(Covariate, levels = c("AGE", "WT", "HT", "CRCL"))) %>% 
  arrange(Covariate, PN) %>%
  select(Covariate, PN, length, median, min, max) %>% 
  # group_by(Covariate, PN) %>% 
  # summarise(N = n,
  #           `Median[Range]` = paste0(median, " [", min, "-" , max, "]")) %>% 
  rename(`Protocol No.` = PN,
         N              = length) %>% 
  flextable() %>%
  merge_v(part = "body") %>%
  merge_h(part = "body") %>%
  theme_box() %>% 
  font(fontname="Times New Roman", part = "all") %>%
  fontsize(size=10, part = "all") %>%
  align(align = "left", part="all") %>% 
  width(width = 0.8)

library(officer)

## word export
doc <- read_docx() %>%
  
  body_add_par(value = "EDA Table", style = "heading 2") %>%
  body_add_flextable(value = nm_stats) %>%
  
  print(doc, target = "1_data_handling/Tables.docx")

```


```{r PK-profiles, eval=FALSE}

# Mean Linear plot for all trials
ggplot(data=nm_data, 
       aes(x=RNTM, y=CONC, group=DOSE, color=factor(DOSE))) + 
  stat_summary() + stat_summary(geom = "line") + 
  facet_wrap(~PN)  + scale_colour_discrete(name="Dose (mg)") + 
  xlab("Time after dose(hr)") + ylab("Plasma concentration (ng/mL)")

# Mean Linear plot for selected trials
ggplot(data=nm_data %>% filter(PN>=8 & PN!=13 | PN <= 2), 
       aes(x=RNTM, y=CONC, group=DOSE, color=factor(DOSE))) + 
  stat_summary() + stat_summary(geom = "line") + 
  facet_wrap(~PN)  + scale_colour_discrete(name="Dose (mg)") + 
  xlab("Time after dose(hr)") + ylab("Plasma concentration (ng/mL)")

ggsave("1_data_handling/EDA_plot.png")

# Mean Semi-log plot for selected trials
ggplot(data=nm_data %>% filter(PN>=8 & PN!=13 | PN <= 2), 
       aes(x=RNTM, y=CONC, group=DOSE, color=factor(DOSE))) + 
  stat_summary() + stat_summary(geom = "line") + 
  facet_wrap(~PN)  + scale_colour_discrete(name="Dose (mg)") + 
  xlab("Time after dose(hr)") + ylab("Plasma concentration (ng/mL)") +
  scale_y_log10()

ggsave("1_data_handling/EDA_plot_sLog.png")

# indiv. Mean plot for selected trials
ggplot(data=nm_data %>% filter(DRUG==2 & PN==19), 
       aes(x=RNTM, y=CONC)) + 
  geom_line(aes(group=ID), colour="darkgray") +
  stat_summary(colour="red") + stat_summary(geom = "line", colour="red") + 
  facet_grid(PERD~DOSE)  + scale_colour_discrete(name="Dose (mg)") + 
  xlab("Time after dose(hr)") + ylab("Plasma concentration (ng/mL)")

ggsave("1_data_handling/indiv_Mean_plot.png", width=8, height = 6)

```


#### End of document