---
title: "Data handling and simple exploratory data analysis"
author: "Masato Fukae and Tomohiro Sasaki"
date: "`r Sys.Date()`"
output: html_document
---

## Data handling

Welcome to **data handling** session!  
Tutors try to provide accurate code. However, incorrect one may be included. Please use it at your own risk.  
This document was built by `r R.Version()$version.string`.  

```{r setup, include=FALSE}
#install.packages("knitr")
#install.packages("tidyverse")
#install.packages("dplyr")

library(knitr)
library(tidyverse)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE)
```

## Set up

Before handling the dataset, set some libraries.

```{r settings}
path <- "F:/Documents/GitHub/R-for-PMx"               # enviroment - 1
path <- "C:/Users/Sasaki.Tomohiro/Desktop/R-for-PMx"  # enviroment - 2
# default path <- "/cloud/project/R-for-PMx-master"
```


## Import data

Let's import nonmem dataset!

```{r import_and_check}
nm_data <- read_csv(paste0(path, "/Data/PSP4-8-748-s012.csv"))  ## csv fileの読み込み

print(head(nm_data, 10))    ## 最初の10行の表示                   

# print(nrow(nm_data))
# print(ncol(nm_data))
# print(names(nm_data))

#　列数, 行数, 列名の表示
nm_data %>% nrow() %>% print()
nm_data %>% ncol() %>% print()
nm_data %>% names() %>% print()

# スプレッドシートでの表示
View(nm_data)
```

## Exploratory data analysis

Let's see the dataset by summary statistics and visualization!

### Generate Summary statistics

```{r Sum-stats}
## group_by, filter, summarize, slice, n, count
# 薬剤ごとの濃度ポイント数（conc>0）の算出
nm_data %>% group_by(DRUG) %>% filter(CONC > 0) %>% summarise(Nobs = n())
# 薬剤ごと, PN(試験)ごとの濃度ポイント数（conc>0）の算出
nm_data %>% group_by(DRUG, PN) %>% filter(CONC > 0) %>% summarise(Nobs = n())

# IDごとに1行のデータを作成
sl_nm_data <- nm_data %>% group_by(ID) %>% slice(1) %>% ungroup()

# ID数の表示
sl_nm_data %>% nrow() %>% print()
# PN(試験)ごとのID数の表示
sl_nm_data %>% count(PN)

sl_nm_data %>% summarise(Number = n(),
                         meanCRCL = mean(CRCL),
                         meanWT = mean(WT),
                         meanAGE = mean(AGE))

sl_nm_data %>% group_by(PN) %>% summarise(Number = n(),
                                          meanCRCL = mean(CRCL) %>% round(1),
                                          meanWT = mean(WT) %>% round(1),
                                          meanAGE = mean(AGE) %>% round(1)) %>% knitr::kable()

sl_nm_data %>% group_by(PN) %>% summarise(Number = n(),
                                          meanCRCL = mean(CRCL) %>% round(1),
                                          meanWT = mean(WT) %>% round(1),
                                          meanAGE = mean(AGE) %>% round(1),
                                          sdCRCL = sd(CRCL) %>% round(1),
                                          sdWT = sd(WT) %>% round(1),
                                          sdAGE = sd(AGE) %>% round(1)) %>% 
  mutate(CRCL = paste(meanCRCL, sdCRCL, sep = " ﾂｱ ")) %>% 
  mutate(WT = paste(meanWT, sdWT, sep = " ﾂｱ ")) %>% 
  mutate(AGE = paste(meanAGE, sdAGE, sep = " ﾂｱ ")) %>%
  select(PN, Number, CRCL, WT, AGE) %>% 
  rename(Protocol_ID = PN) %>% 
  knitr::kable(align = "c")

```

### Visualize data

```{r}
p <- ggplot(sl_nm_data %>% select(PN, CRCL, WT, AGE))
p <- p + geom_point(aes(x = WT, y = CRCL))
p

p <- p + facet_wrap(~ PN)
p

p <- ggplot(sl_nm_data %>% select(PN, CRCL, WT, AGE, HLTH))
p <- p + geom_point(aes(x = WT, y = CRCL, fill = factor(HLTH)), shape = 21, alpha = 0.5)
p <- p + theme_bw()
p <- p + labs(x = "Body weight (kg)", y = "Creatine clearance (mL/min)", fill = "Healthy volunteer")
p <- p + scale_x_continuous(limits = c(25, NA))
p <- p + scale_y_continuous(limits = c(0, NA))
p <- p + theme(legend.position = "top")
p

```

#### End of document